---
title: "pspforecast readme"
output: github_document
---


# pspforecast

Shellfish toxicity forecast serving package


## Requirements

  + [R v4+](https://www.r-project.org/)
  
  + [rlang](https://CRAN.R-project.org/package=rlang)
  
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  
  + [readr](https://CRAN.R-project.org/package=readr)
  
  + [tidyr](https://CRAN.R-project.org/package=tidyr)

## Installation

```
remotes::install_github("BigelowLab/pspforecast")
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)

library(pspdata)
library(pspforecast)
```

## Reading the forecast database 

### Variables:
 + version - the version/configuration of the model used to make the prediction
 + ensemble_n - number of ensemble members used to generate prediction
 
 + location - the sampling station the forecast is for
 
 + date - the date the forecast was made on
 
 + name - site name
 
 + lat - latitude
 
 + lon - longitude
 
 + class_bins - the bins used to classify shellfish total toxicity (i.e. 0: 0-10, 1: 10-30, 2: 30-80, 3: >80)
 
 + forecast_date - the date the forecast is valid for (i.e. one week ahead of when it was made)
 
 + predicted_class - the predicted classification at the location listed on the forecast_date (in this case 0-3)
 + p_0 - class 0 probability
 + p_1 - class 1 probability
 + p_2 - class 2 probability
 + p_3 - class 3 probability
 + p3_sd - class 3 probability standard deviation
 + p_3_min - class 3 minimum probability (from ensemble run)
 + p_3_max - class 3 maximum probability (from ensemble run)
 + predicted_class - the predicted classification


```{r echo=FALSE}
psp <- read_psp_data() %>% 
  dplyr::mutate(year = as.numeric(format(date, format="%Y")),
                doy = as.numeric(format(date, format="%j")),
                week = as.numeric(format(date, format="%U")))
```


```{r}
predictions <- pspforecast::read_forecast()

glimpse(predictions)
```


```{r echo=FALSE}
pred_w_results <- add_forecast_results(predictions, toxin_measurements = psp)
```

## 2022 Season Performance

### Confusion Matrix

```{r echo=FALSE}
num_levels <- 4
levels <- seq(from=0, to=(num_levels-1))


cm <- as.data.frame(table(predicted = factor(pred_w_results$predicted_class, levels), actual = factor(pred_w_results$class, levels))) %>% 
  mutate(frac = round(Freq/sum(Freq)*100)) %>% 
  mutate(frac = sapply(.data$frac, function(x) if (x == "0") {x = "<1"} else {x}))
  #mutate(frac = sapply(.data$frac, function(x) if (x == "0") {x = "<1"} else {x}))

confusion_matrix <- ggplot2::ggplot(data = cm,
                                    mapping = ggplot2::aes(x = .data$predicted, y = .data$actual)) +
  ggplot2::geom_tile(ggplot2::aes(fill = log(.data$Freq+1))) +
  ggplot2::geom_text(ggplot2::aes(label = sprintf("%1.0f", .data$Freq)), size=8) +
  #ggplot2::geom_text(ggplot2::aes(label = sprintf("%1.0f%%", .data$frac)), vjust = -1, hjust=1, size=5) +
  ggplot2::geom_text(ggplot2::aes(label = paste(.data$frac, "%", sep="")), vjust = 3, size=4)+
  ggplot2::scale_fill_gradient(low = "white", 
                               high = "blue") +
  ggplot2::labs(x = "Predicted Classifications", 
                y = "Actual Classifications") +
  ggplot2::theme_linedraw() +
  ggplot2::theme(axis.text=  ggplot2::element_text(size=14),
                 axis.title= ggplot2::element_text(size=14,face="bold"),
                 title =     ggplot2::element_text(size = 14, face = "bold"),
                 legend.position = "none")

confusion_matrix

```

### Probability of Closure-level Toxicity vs Measured Toxicity

```{r echo=FALSE}

p3_v_tox <- ggplot2::ggplot(data = pred_w_results, ggplot2::aes(x=.data$p_3, y=.data$toxicity)) +
  ggplot2::geom_point(alpha=0.7, size=3) +
  ggplot2::labs(x = "Predicted Probability of Closure-level Toxicity",
                y = "Measured Toxicity") +
  ggplot2::geom_hline(yintercept=80, linetype="dashed")

p3_v_tox
```


### Last Updated

```{r echo=FALSE}
Sys.Date()
```

