---
title: "pspforecast"
output: github_document
---


Shellfish toxicity (PSP) forecast serving package

For the current 2026 Maine PSP predictions, click [here](https://github.com/BigelowLab/pspforecast/blob/master/inst/forecastdb/dmr_webpage_table.csv)

For past season results, check out this [shiny application](https://eco.bigelow.org/pspforecast/)


## Requirements

  + [R v4+](https://www.r-project.org/)
  
  + [rlang](https://CRAN.R-project.org/package=rlang)
  
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  
  + [readr](https://CRAN.R-project.org/package=readr)
  
  + [tidyr](https://CRAN.R-project.org/package=tidyr)
  
  + [httr](https://CRAN.R-project.org/package=httr)

## Installation

```
remotes::install_github("BigelowLab/pspforecast")
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(plotly)
  
  library(pspdata)
  library(pspforecast)
  
  library(patchwork)
})
```

## Reading the forecast database 

### Variables:
 + version - the version/configuration of the model used to make the prediction
 
 + ensemble_n - number of ensemble members used to generate prediction
 
 + location - the sampling station the forecast is for
 
 + date - the date the forecast was made on
 
 + name - site name
 
 + lat - latitude
 
 + lon - longitude
 
 + class_bins - the bins used to classify shellfish total toxicity (i.e. 0: 0-10, 1: 10-30, 2: 30-80, 3: >80)
 
 + forecast_date - the date the forecast is valid for (i.e. one week ahead of when it was made)
 
 + predicted_class - the predicted classification at the location listed on the forecast_date (in this case 0-3)
 + p_0 - class 0 probability
 + p_1 - class 1 probability
 + p_2 - class 2 probability
 + p_3 - class 3 probability
 + p3_sd - class 3 probability standard deviation
 + p_3_min - class 3 minimum probability (from ensemble run)
 + p_3_max - class 3 maximum probability (from ensemble run)
 + predicted_class - the predicted classification
 
### Metrics:
 
 + tp - The model predicted class 3 and the following week's measurement was class 3
 + fp - The model predicted class 3 and the following week's measurement was not class 3
 + tn - The model predicted class 0,1,2 and the following week's measurement was in class 0,1,2
 + fn - The model predicted class 0,1,2 and the following week's measurement was class 3
 + accuracy - Measure of how many correct classifications were predicted
 + cl_accuracy - Considering predictions are those that correctly predicted toxicity above or below the closure limit or not
 + precision - TP/(TP+FP)
 + sensitivity - TP/(TP+FN)
 + specificity - TN/(TN+FP)
 + f_1 - (2TP)/(2TP-FP+FN)


```{r echo=FALSE, message=FALSE, warning=FALSE}
psp <- read_psp_data()
```

## Overall Results (2021-2025 seasons combined)


```{r echo=FALSE}
all_results <- read_all_results() |>
  distinct()
```

Predictions evaluated:

```{r echo=FALSE}
nrow(all_results)
```

Metrics:

```{r echo=FALSE}
forecast_metrics(all_results, measured_col = "class", positive_col = "p_3")
```


## 2026 (current season) Results


```{r echo=FALSE}
predictions <- read_forecast(year = "2026") |>
  mutate(species = "mytilus")
```


```{r echo=FALSE}
pred_w_results <- add_forecast_results(predictions, toxin_measurements = psp) 
```


```{r cm26, echo=FALSE, cahce=FALSE}
num_levels <- 4
levels <- seq(from=0, to=(num_levels-1))

cm <- as.data.frame(table(predicted = factor(pred_w_results$predicted_class, levels), 
                          actual = factor(pred_w_results$class, levels))) |> 
  mutate(outcome = as.factor(case_when(predicted %in% c(0,1,2) & actual %in% c(0,1,2) ~"TN",
                                       predicted == 3 & actual == 3 ~"TP",
                                       predicted %in% c(0,1,2) & actual == 3 ~ "FN",
                                       predicted == 3 & actual %in% c(0,1,2) ~ "FP")))

ggplot2::ggplot(data = cm, ggplot2::aes(x=.data$predicted, y=.data$actual)) +
  ggplot2::geom_tile(aes(fill = outcome)) +
  ggplot2::geom_text(ggplot2::aes(label = Freq), size=7, fontface="bold") +
  scale_fill_discrete(palette = c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c")) +
  ggplot2::labs(x = "Predicted Classifications", 
                y = "Actual Classifications") +
  ggplot2::theme_linedraw() +
  ggplot2::theme(axis.text=  ggplot2::element_text(size=14),
                 axis.title= ggplot2::element_text(size=18,face="bold"),
                 title =     ggplot2::element_text(size = 14, face = "bold"),
                 legend.position = "bottom",
                 legend.title = element_blank(),
                 legend.text = element_text(size = 14, face="bold"),
                 strip.text.x = element_text(size = 20, face="bold"))
```


```{r scatter26, echo=FALSE, cache=FALSE}
p3_v_tox_26 <- ggplot2::ggplot(data = pred_w_results, ggplot2::aes(x=.data$p_3, y=.data$toxicity)) +
  ggplot2::geom_point(alpha=0.7, size=3) +
  ggplot2::labs(x = "Predicted Probability (%) of Closure-level Toxicity",
                y = "Measured Toxicity") +
  ggplot2::geom_hline(yintercept=80, linetype="dashed") +
  ggplot2::xlim(0,100)

p3_v_tox_26
```

### Predictions evaluated

```{r echo=FALSE}
nrow(pred_w_results)
```


### Metrics 


```{r echo=FALSE}
forecast_metrics(pred_w_results, measured_col = "class", positive_col = "p_3")
```


### Last Updated

```{r echo=FALSE}
Sys.Date()
```

